\documentclass[main.tex]{subfiles}

\begin{document}
\chapter{Validation}
\label{chap:validation}
In order to validate the implementation of BFAST and BFAST0n, four datasets were
chosen, each testing a specific property of the algorithm. Three of the datasets were
taken from the paper by Verbesselt et al. \cite{bfast} and one from the standard
dataset library for the R programming language \cite{r-datasets}

The implementation was tested using the same methodology as in the paper by
Verbesselt et al. \cite{bfast}. The \texttt{bfast} and \texttt{bfast0n}
functions were applied to the simulated 16-day Normalized Difference Vegetation
Index(NDVI) time series, 16-day NDVI Moderate Resolution Imaging
Spectroradiometer (MODIS) composites and others. The results were compared to
the same data applied to the original R-implementation of BFAST and BFAST0n
\cite{bfast-github}.\\\\ 
Following list datasets were used:
\begin{itemize}
\item \texttt{harvest}
  \begin{itemize}
  \item \textbf{Description:}\\
  16-day NDVI Time Series for a Pinus Radiata Plantation, with approximately 23
  observation per year.
  \item \textbf{Source:} \cite{bfast}
  \end{itemize}
\item \texttt{modisraster}
  \begin{itemize}
  \item \textbf{Description:}\\ A raster brick of 16-day satellite image NDVI
    time series for a small subset in south eastern Somalia.
  \item \textbf{Source:} \cite{bfast_monitor1}
  \end{itemize}
\item \texttt{ndvi}
  \begin{itemize}
  \item \textbf{Description:}\\ A random NDVI time series. Frequency is set to 24
  \end{itemize}
\item \texttt{simts}
  \begin{itemize}
  \item \textbf{Description:}\\ Simulated seasonal 16-day NDVI time series.
  \item \textbf{Source:} \cite{bfast}
  \end{itemize}
\end{itemize}

\section{\texttt{nile}}
\label{sec:val_nile}

\section{\texttt{harvest}}
\label{sec:val_harvest}

\section{\texttt{simts}}
\label{sec:val_simts}

\section{\texttt{ndvi}}
\label{sec:val_ndvi}



\biblio
\end{document}
